---
title: "STAT535_Final_Project"
author: "Yecheng Cao, Qi Chen, Travis Xiaoyu Zhao"
date: "11/28/2021"
output: html_document
---
* [Introduction]
  * [Objectives]
  * [Web Scraping]
* [Prediction]
  * [Ordinary Least Squares]

  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

```{r, echo=FALSE}
set.seed(4321)
```

# Introduction {-}

## Objectives ############################# 

<br>
**Based on the English Premier League Tables and all Game Day results in past 5 seasons (2016-17 season until 2020-21 season), we would like to use a few techniques to predict the champion in this season 2021-22.**
<br>

<br>
**We scraped the data from myfootballfacts.com and fixturedownload.com/sport/football. Here are the attached original htmls. (https://www.myfootballfacts.com/premier-league/all-time-premier-league/premier-league-ladders/) and (https://fixturedownload.com/sport/football)**
<br>



## Web Scraping ############################# 

```{r, echo=TRUE}
# Read the raw data
tables_html = readLines("/Users/chenqi/Desktop/535 final/YechengCao/Tables.html", warn = FALSE)

# Seasons
seasons = tables_html[grep("<td class=\"dtr1 dtc", tables_html)][1:24]

# Champion teams with corresponding points
rank_1 = tables_html[grep("<td class=\"dtr2 dtc", tables_html)][1:24]

# Runner-ups teams
rank_2 = tables_html[grep("<td class=\"dtr3 dtc", tables_html)][1:24]

# All rankings below with same technique
rank_3 = tables_html[grep("<td class=\"dtr4 dtc", tables_html)][1:24]
rank_4 = tables_html[grep("<td class=\"dtr5 dtc", tables_html)][1:24]
rank_5 = tables_html[grep("<td class=\"dtr6 dtc", tables_html)][1:24]
rank_6 = tables_html[grep("<td class=\"dtr7 dtc", tables_html)][1:24]
rank_7 = tables_html[grep("<td class=\"dtr8 dtc", tables_html)][1:24]
rank_8 = tables_html[grep("<td class=\"dtr9 dtc", tables_html)][1:24]
rank_9 = tables_html[grep("<td class=\"dtr10 dtc", tables_html)][1:24]
rank_10 = tables_html[grep("<td class=\"dtr11 dtc", tables_html)][1:24]
rank_11 = tables_html[grep("<td class=\"dtr12 dtc", tables_html)][1:24]
rank_12 = tables_html[grep("<td class=\"dtr13 dtc", tables_html)][1:24]
rank_13 = tables_html[grep("<td class=\"dtr14 dtc", tables_html)][1:24]
rank_14 = tables_html[grep("<td class=\"dtr15 dtc", tables_html)][1:24]
rank_15 = tables_html[grep("<td class=\"dtr16 dtc", tables_html)][1:24]
rank_16 = tables_html[grep("<td class=\"dtr17 dtc", tables_html)][1:24]
rank_17 = tables_html[grep("<td class=\"dtr18 dtc", tables_html)][1:24]
rank_18 = tables_html[grep("<td class=\"dtr19 dtc", tables_html)][1:24]
rank_19 = tables_html[grep("<td class=\"dtr20 dtc", tables_html)][1:24]
rank_20 = tables_html[grep("<td class=\"dtr21 dtc", tables_html)][1:24]
```

```{r, echo=TRUE}
data_processing_lead = function(x){
  temp = vector()
  for(i in 1:length(x)){
    if(sum(nchar(x[i]))>28){ #if there are more than 28 chars, those lines are either team names or pts that we want
      temp[i] = x[i]
    }
    temp = na.exclude(temp) # remove all NAs
  }
  pattern = "(.*)\">|</td>"
  temp = gsub(pattern, "", temp) # remove useless strings
  done = temp[1:length(temp)] # keep the information we want
  return(done)
}
rank_1_clear = data_processing_lead(rank_1)
rank_2_clear = data_processing_lead(rank_2)
rank_3_clear = data_processing_lead(rank_3)
rank_4_clear = data_processing_lead(rank_4)
rank_5_clear = data_processing_lead(rank_5)
rank_6_clear = data_processing_lead(rank_6)
rank_7_clear = data_processing_lead(rank_7)
rank_8_clear = data_processing_lead(rank_8)

data_processing_lag = function(x){ # the only difference is the "if" statement because the lines start from rank 9 teams are longer, i.e. dtr10 and so on have 2 digits 
  temp = vector()
  for(i in 1:length(x)){
    if(sum(nchar(x[i]))>29){ #if there are more than 29 chars, those lines are either team names or pts that we want
      temp[i] = x[i]
    }
    temp = na.exclude(temp) # remove all NAs
  }
  pattern = "(.*)\">|</td>"
  temp = gsub(pattern, "", temp) # remove useless strings
  done = temp[1:length(temp)] # keep the information we want
  return(done)
}
seasons_clear = data_processing_lag(seasons)
rank_9_clear = data_processing_lag(rank_9)
rank_10_clear = data_processing_lag(rank_10)
rank_11_clear = data_processing_lag(rank_11)
rank_12_clear = data_processing_lag(rank_12)
rank_13_clear = data_processing_lag(rank_13)
rank_14_clear = data_processing_lag(rank_14)
rank_15_clear = data_processing_lag(rank_15)
rank_16_clear = data_processing_lag(rank_16)
rank_17_clear = data_processing_lag(rank_17)
rank_18_clear = data_processing_lag(rank_18)
rank_19_clear = data_processing_lag(rank_19)
rank_20_clear = data_processing_lag(rank_20)

# Remove the first element that represents the ranking starting from rank 10 teams
rank_10_clear = rank_10_clear[-1]
rank_11_clear = rank_11_clear[-1]
rank_12_clear = rank_12_clear[-1]
rank_13_clear = rank_13_clear[-1]
rank_14_clear = rank_14_clear[-1]
rank_15_clear = rank_15_clear[-1]
rank_16_clear = rank_16_clear[-1]
rank_17_clear = rank_17_clear[-1]
rank_18_clear = rank_18_clear[-1]
rank_19_clear = rank_19_clear[-1]
rank_20_clear = rank_20_clear[-1]

# Check all vectors are in correct length so that we could put them into a data frame
all.equal(
  length(seasons_clear),
  length(rank_1_clear),
  length(rank_2_clear),
  length(rank_3_clear),
  length(rank_4_clear),
  length(rank_5_clear),
  length(rank_6_clear),
  length(rank_7_clear),
  length(rank_8_clear),
  length(rank_9_clear),
  length(rank_10_clear),
  length(rank_11_clear),
  length(rank_12_clear),
  length(rank_13_clear),
  length(rank_14_clear),
  length(rank_15_clear),
  length(rank_16_clear),
  length(rank_17_clear),
  length(rank_18_clear),
  length(rank_19_clear),
  length(rank_20_clear))
```


```{r, echo=TRUE}
# Combine all previous info together into a data frame
tables = rbind(seasons_clear, 
               rank_1_clear, 
               rank_2_clear,
               rank_3_clear,
               rank_4_clear,
               rank_5_clear,
               rank_6_clear,
               rank_7_clear,
               rank_8_clear,
               rank_9_clear,
               rank_10_clear,
               rank_11_clear,
               rank_12_clear,
               rank_13_clear,
               rank_14_clear,
               rank_15_clear,
               rank_16_clear,
               rank_17_clear,
               rank_18_clear,
               rank_19_clear,
               rank_20_clear)

colnames(tables) = c("2013-14", "Pts", "2014-15", "Pts", "2015-16", "Pts", "2016-17", "Pts", "2017-18", "Pts",
                     "2018-19", "Pts", "2019-20", "Pts", "2020-21", "Pts")

rownames(tables) = c("Season", 1:20)

# Table for Seasons 2013-14 until 2020-21
head(tables)
```


```{r, echo=TRUE}
# Read the raw data for matches
result_2021_html = readLines("/Users/chenqi/Desktop/535 final/YechengCao/2020_21.html", warn = FALSE)
result_1920_html = readLines("/Users/chenqi/Desktop/535 final/YechengCao/2019_20.html", warn = FALSE)
result_1819_html = readLines("/Users/chenqi/Desktop/535 final/YechengCao/2018_19.html", warn = FALSE)
result_1718_html = readLines("/Users/chenqi/Desktop/535 final/YechengCao/2017_18.html", warn = FALSE)
result_1617_html = readLines("/Users/chenqi/Desktop/535 final/YechengCao/2016_17.html", warn = FALSE)
```

```{r, echo=TRUE}
teams_processing = function(x){
  temp = x[grep("                        <td>[A-Z]|                        <td>bet", x)]
  pattern = "(.*)<td>|</td>"
  temp_2 = gsub(pattern, "", temp) # remove useless strings
  done = vector()
  for(i in 1:length(temp_2)){
    if((i %% 3)!=1){
      done[i] = temp_2[i]
    }
  }
  done = na.exclude(done) # remove all NAs
  done = done[1:760]
  return(done)
}

goals_processing = function(x){
  temp = x[grep("                                                    <td>\\d+", x)]
  pattern = "(.*)<td>|</td>"
  temp_2 = gsub(pattern, "", temp) # remove useless strings
  home = vector()
  away = vector()
  result = vector()
  for(i in 1:length(temp_2)){
    home[i] = substr(temp_2[i], 1, 1)
    away[i] = substr(temp_2[1], 5, 5)
    result[2*i-1] = home[i]
    result[2*i] = away[i]
  }
  done = na.exclude(result) # remove all NAs
  return(done)
}
```

```{r, echo=TRUE}
matches = cbind(teams_processing(result_1617_html), goals_processing(result_1617_html),
                teams_processing(result_1718_html), goals_processing(result_1718_html),
                teams_processing(result_1819_html), goals_processing(result_1819_html),
                teams_processing(result_1920_html), goals_processing(result_1920_html),
                teams_processing(result_2021_html), goals_processing(result_2021_html))

colnames(matches) = c("2016-17", "Goal(s)", "2017-18", "Goal(s)", "2018-19", "Goal(s)",
                      "2019-20", "Goal(s)", "2020-21", "Goal(s)")

rownames(matches) = rep(c("Home", "Away"), 380)

# All match results for Seasons 2016-17 until 2020-21
head(matches)
```



# Prediction {-}

## Ordinary Least Squares ############################# 

