---
title: "STAT535_Final_Project"
author: "Yecheng Cao, Qi Chen, Travis Xiaoyu Zhao"
date: "11/28/2021"
output: html_document
---
* [Introduction]
  * [Objectives]
  * [Web Scraping]
* [Prediction]
  * [Ordinary Least Squares]

  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vcd)
rm(list = ls())
```

```{r, echo=FALSE}
set.seed(4321)
```

# Introduction {-}

## Objectives ############################# 

<br>
**Based on the English Premier League Tables and all Game Day results in past 5 seasons (2016-17 season until 2020-21 season), we would like to use a few techniques to predict the champion in this season 2021-22.**
<br>

<br>
**We scraped the data from myfootballfacts.com and fixturedownload.com/sport/football. Here are the attached original htmls. (https://www.myfootballfacts.com/premier-league/all-time-premier-league/premier-league-ladders/) and (https://fixturedownload.com/sport/football)**
<br>



## Web Scraping ############################# 

```{r, echo=TRUE}
# Read the raw data
#tables_html = readLines("/Users/yechengcao/Desktop/535/Final/Tables.html", warn = FALSE)
tables_html = readLines("Tables.html", warn = FALSE)

# Seasons
seasons = tables_html[grep("<td class=\"dtr1 dtc", tables_html)][1:24]

# Champion teams with corresponding points
rank_1 = tables_html[grep("<td class=\"dtr2 dtc", tables_html)][1:24]

# Runner-ups teams
rank_2 = tables_html[grep("<td class=\"dtr3 dtc", tables_html)][1:24]

# All rankings below with same technique
rank_3 = tables_html[grep("<td class=\"dtr4 dtc", tables_html)][1:24]
rank_4 = tables_html[grep("<td class=\"dtr5 dtc", tables_html)][1:24]
rank_5 = tables_html[grep("<td class=\"dtr6 dtc", tables_html)][1:24]
rank_6 = tables_html[grep("<td class=\"dtr7 dtc", tables_html)][1:24]
rank_7 = tables_html[grep("<td class=\"dtr8 dtc", tables_html)][1:24]
rank_8 = tables_html[grep("<td class=\"dtr9 dtc", tables_html)][1:24]
rank_9 = tables_html[grep("<td class=\"dtr10 dtc", tables_html)][1:24]
rank_10 = tables_html[grep("<td class=\"dtr11 dtc", tables_html)][1:24]
rank_11 = tables_html[grep("<td class=\"dtr12 dtc", tables_html)][1:24]
rank_12 = tables_html[grep("<td class=\"dtr13 dtc", tables_html)][1:24]
rank_13 = tables_html[grep("<td class=\"dtr14 dtc", tables_html)][1:24]
rank_14 = tables_html[grep("<td class=\"dtr15 dtc", tables_html)][1:24]
rank_15 = tables_html[grep("<td class=\"dtr16 dtc", tables_html)][1:24]
rank_16 = tables_html[grep("<td class=\"dtr17 dtc", tables_html)][1:24]
rank_17 = tables_html[grep("<td class=\"dtr18 dtc", tables_html)][1:24]
rank_18 = tables_html[grep("<td class=\"dtr19 dtc", tables_html)][1:24]
rank_19 = tables_html[grep("<td class=\"dtr20 dtc", tables_html)][1:24]
rank_20 = tables_html[grep("<td class=\"dtr21 dtc", tables_html)][1:24]

#rank <- rep(NA, 20)
#for (i in 1:20){
#  rank[i] <- tables_html[grep(paste0("<td class=\"dtr", i+1, " dtc"), tables_html)][1:24]
#}

#rank_1 == rank[1]
```

```{r, echo=TRUE}
data_processing_lead = function(x){
  temp = vector()
  for(i in 1:length(x)){
    if(sum(nchar(x[i]))>28){ #if there are more than 28 chars, those lines are either team names or pts that we want
      temp[i] = x[i]
    }
    temp = na.exclude(temp) # remove all NAs
  }
  pattern = "(.*)\">|</td>"
  temp = gsub(pattern, "", temp) # remove useless strings
  done = temp[1:length(temp)] # keep the information we want
  return(done)
}
rank_1_clear = data_processing_lead(rank_1)
rank_2_clear = data_processing_lead(rank_2)
rank_3_clear = data_processing_lead(rank_3)
rank_4_clear = data_processing_lead(rank_4)
rank_5_clear = data_processing_lead(rank_5)
rank_6_clear = data_processing_lead(rank_6)
rank_7_clear = data_processing_lead(rank_7)
rank_8_clear = data_processing_lead(rank_8)

data_processing_lag = function(x){ # the only difference is the "if" statement because the lines start from rank 9 teams are longer, i.e. dtr10 and so on have 2 digits 
  temp = vector()
  for(i in 1:length(x)){
    if(sum(nchar(x[i]))>29){ #if there are more than 29 chars, those lines are either team names or pts that we want
      temp[i] = x[i]
    }
    temp = na.exclude(temp) # remove all NAs
  }
  pattern = "(.*)\">|</td>"
  temp = gsub(pattern, "", temp) # remove useless strings
  done = temp[1:length(temp)] # keep the information we want
  return(done)
}
seasons_clear = data_processing_lag(seasons)
rank_9_clear = data_processing_lag(rank_9)
rank_10_clear = data_processing_lag(rank_10)
rank_11_clear = data_processing_lag(rank_11)
rank_12_clear = data_processing_lag(rank_12)
rank_13_clear = data_processing_lag(rank_13)
rank_14_clear = data_processing_lag(rank_14)
rank_15_clear = data_processing_lag(rank_15)
rank_16_clear = data_processing_lag(rank_16)
rank_17_clear = data_processing_lag(rank_17)
rank_18_clear = data_processing_lag(rank_18)
rank_19_clear = data_processing_lag(rank_19)
rank_20_clear = data_processing_lag(rank_20)

# Remove the first element that represents the ranking starting from rank 10 teams
rank_10_clear = rank_10_clear[-1]
rank_11_clear = rank_11_clear[-1]
rank_12_clear = rank_12_clear[-1]
rank_13_clear = rank_13_clear[-1]
rank_14_clear = rank_14_clear[-1]
rank_15_clear = rank_15_clear[-1]
rank_16_clear = rank_16_clear[-1]
rank_17_clear = rank_17_clear[-1]
rank_18_clear = rank_18_clear[-1]
rank_19_clear = rank_19_clear[-1]
rank_20_clear = rank_20_clear[-1]

# Check all vectors are in correct length so that we could put them into a data frame
all.equal(
  length(seasons_clear),
  length(rank_1_clear),
  length(rank_2_clear),
  length(rank_3_clear),
  length(rank_4_clear),
  length(rank_5_clear),
  length(rank_6_clear),
  length(rank_7_clear),
  length(rank_8_clear),
  length(rank_9_clear),
  length(rank_10_clear),
  length(rank_11_clear),
  length(rank_12_clear),
  length(rank_13_clear),
  length(rank_14_clear),
  length(rank_15_clear),
  length(rank_16_clear),
  length(rank_17_clear),
  length(rank_18_clear),
  length(rank_19_clear),
  length(rank_20_clear))
```


```{r, echo=TRUE}
# Combine all previous info together into a data frame
tables = rbind(seasons_clear, 
               rank_1_clear, 
               rank_2_clear,
               rank_3_clear,
               rank_4_clear,
               rank_5_clear,
               rank_6_clear,
               rank_7_clear,
               rank_8_clear,
               rank_9_clear,
               rank_10_clear,
               rank_11_clear,
               rank_12_clear,
               rank_13_clear,
               rank_14_clear,
               rank_15_clear,
               rank_16_clear,
               rank_17_clear,
               rank_18_clear,
               rank_19_clear,
               rank_20_clear)

colnames(tables) = c("2013-14", "Pts", "2014-15", "Pts", "2015-16", "Pts", "2016-17", "Pts", "2017-18", "Pts",
                     "2018-19", "Pts", "2019-20", "Pts", "2020-21", "Pts")

rownames(tables) = c("Season", 1:20)

# Table for Seasons 2013-14 until 2020-21
head(tables)
```


```{r, echo=TRUE}
# Read the raw data for matches
#result_2021_html = readLines("/Users/yechengcao/Desktop/535/Final/2020_21.html", warn = FALSE)
#result_1920_html = readLines("/Users/yechengcao/Desktop/535/Final/2019_20.html", warn = FALSE)
#result_1819_html = readLines("/Users/yechengcao/Desktop/535/Final/2018_19.html", warn = FALSE)
#result_1718_html = readLines("/Users/yechengcao/Desktop/535/Final/2017_18.html", warn = FALSE)
#result_1617_html = readLines("/Users/yechengcao/Desktop/535/Final/2016_17.html", warn = FALSE)

result_2021_html = readLines("2020_21.html", warn = FALSE)
result_1920_html = readLines("2019_20.html", warn = FALSE)
result_1819_html = readLines("2018_19.html", warn = FALSE)
result_1718_html = readLines("2017_18.html", warn = FALSE)
result_1617_html = readLines("2016_17.html", warn = FALSE)
```

```{r, echo=TRUE}
teams_processing = function(x){
  temp = x[grep("                        <td>[A-Z]|                        <td>bet", x)]
  pattern = "(.*)<td>|</td>"
  temp_2 = gsub(pattern, "", temp) # remove useless strings
  done = vector()
  for(i in 1:length(temp_2)){
    if((i %% 3)!=1){
      done[i] = temp_2[i]
    }
  }
  done = na.exclude(done) # remove all NAs
  done = done[1:760]
  return(done)
}

goals_processing = function(x){
  temp = x[grep("                                                    <td>\\d+", x)]
  pattern = "(.*)<td>|</td>"
  temp_2 = gsub(pattern, "", temp) # remove useless strings
  home = vector()
  away = vector()
  result = vector()
  for(i in 1:length(temp_2)){
    home[i] = substr(temp_2[i], 1, 1)
    away[i] = substr(temp_2[i], 5, 5)
    result[2*i-1] = home[i]
    result[2*i] = away[i]
  }
  done = na.exclude(result) # remove all NAs
  return(done)
}

```

```{r, echo=TRUE}
matches = cbind(teams_processing(result_1617_html), goals_processing(result_1617_html),
                teams_processing(result_1718_html), goals_processing(result_1718_html),
                teams_processing(result_1819_html), goals_processing(result_1819_html),
                teams_processing(result_1920_html), goals_processing(result_1920_html),
                teams_processing(result_2021_html), goals_processing(result_2021_html))

#colnames(matches) = c("2016-17", "Goal(s)", "2017-18", "Goal(s)", "2018-19", "Goal(s)",
#                      "2019-20", "Goal(s)", "2020-21", "Goal(s)")

colnames(matches) = c("s1617", "g1617", "s1718", "g1718", "s1819", "g1819",
                      "s1920", "g1920", "s2021", "g2021")
rownames(matches) = rep(c("Home", "Away"), 380)


# All match results for Seasons 2016-17 until 2020-21
head(matches)
```



# Prediction {-}
```{r}
results_1617 <- data.frame(matrix(NA, nrow = 380, ncol = 5))
for(i in 1:760){
  if((i %% 2)!=0){
    results_1617[(i%/%2)+1,1] = matches[i,1]
    results_1617[(i%/%2)+1,3] = matches[i,2]
  } else {
    results_1617[i/2,2] = matches[i,1]
    results_1617[i/2,4] = matches[i,2]
    
    if(results_1617[i/2,3]>results_1617[i/2,4]){
      results_1617[i/2,5] = "H"
    } else if(results_1617[i/2,3]<results_1617[i/2,4]){
      results_1617[i/2,5] = "A"
    } else{
      results_1617[i/2,5] = "D"
    }
  }
}
colnames(results_1617) = c("HomeTeam", "AwayTeam", "FTHG", "FTAG", "FTR")

results_1718 <- data.frame(matrix(NA, nrow = 380, ncol = 5))
for(i in 1:760){
  if((i %% 2)!=0){
    results_1718[(i%/%2)+1,1] = matches[i,3]
    results_1718[(i%/%2)+1,3] = matches[i,4]
  } else {
    results_1718[i/2,2] = matches[i,3]
    results_1718[i/2,4] = matches[i,4]
    
    if(results_1718[i/2,3]>results_1718[i/2,4]){
      results_1718[i/2,5] = "H"
    } else if(results_1718[i/2,3]<results_1718[i/2,4]){
      results_1718[i/2,5] = "A"
    } else{
      results_1718[i/2,5] = "D"
    }
  }
}
colnames(results_1718) = c("HomeTeam", "AwayTeam", "FTHG", "FTAG", "FTR")

results_1819 <- data.frame(matrix(NA, nrow = 380, ncol = 5))
for(i in 1:760){
  if((i %% 2)!=0){
    results_1819[(i%/%2)+1,1] = matches[i,5]
    results_1819[(i%/%2)+1,3] = matches[i,6]
  } else {
    results_1819[i/2,2] = matches[i,5]
    results_1819[i/2,4] = matches[i,6]
    
    if(results_1819[i/2,3]>results_1819[i/2,4]){
      results_1819[i/2,5] = "H"
    } else if(results_1819[i/2,3]<results_1819[i/2,4]){
      results_1819[i/2,5] = "A"
    } else{
      results_1819[i/2,5] = "D"
    }
  }
}
colnames(results_1819) = c("HomeTeam", "AwayTeam", "FTHG", "FTAG", "FTR")

results_1920 <- data.frame(matrix(NA, nrow = 380, ncol = 5))
for(i in 1:760){
  if((i %% 2)!=0){
    results_1920[(i%/%2)+1,1] = matches[i,7]
    results_1920[(i%/%2)+1,3] = matches[i,8]
  } else {
    results_1920[i/2,2] = matches[i,7]
    results_1920[i/2,4] = matches[i,8]
    
    if(results_1920[i/2,3]>results_1920[i/2,4]){
      results_1920[i/2,5] = "H"
    } else if(results_1920[i/2,3]<results_1920[i/2,4]){
      results_1920[i/2,5] = "A"
    } else{
      results_1920[i/2,5] = "D"
    }
  }
}
colnames(results_1920) = c("HomeTeam", "AwayTeam", "FTHG", "FTAG", "FTR")

results_2021 <- data.frame(matrix(NA, nrow = 380, ncol = 5))
for(i in 1:760){
  if((i %% 2)!=0){
    results_2021[(i%/%2)+1,1] = matches[i,9]
    results_2021[(i%/%2)+1,3] = matches[i,10]
  } else {
    results_2021[i/2,2] = matches[i,9]
    results_2021[i/2,4] = matches[i,10]
    
    if(results_2021[i/2,3]>results_2021[i/2,4]){
      results_2021[i/2,5] = "H"
    } else if(results_2021[i/2,3]<results_2021[i/2,4]){
      results_2021[i/2,5] = "A"
    } else{
      results_2021[i/2,5] = "D"
    }
  }
}
colnames(results_2021) = c("HomeTeam", "AwayTeam", "FTHG", "FTAG", "FTR")
```


```{r}
# Temporarily use csv data
# test_data <- rbind(read_csv("season-1617_csv.csv")[2:7], read_csv("season-1718_csv.csv")[2:7], read_csv("season-1819_csv.csv")[2:7])
test_data = rbind(results_1617, results_1718, results_1819, results_1920, results_2021)
```

**Data Summary**
```{r}
# Summarize home power
HomePower <- test_data %>% 
  group_by(HomeTeam) %>% 
  summarize(HWins = sum(FTR == "H"), HDraws = sum(FTR == "D"), HLoses = 19*3 - HWins - HDraws,
            HP = 3 * HWins + 1 * HDraws, HPower = HP / (19*3*3))

# Summarize away power
AwayPower <- test_data %>% 
  group_by(AwayTeam) %>% 
  summarize(AWins = sum(FTR == "A"), ADraws = sum(FTR == "D"), ALoses = 19*3 - AWins - ADraws,
            AP = 3 * AWins + 1 * ADraws, APower = AP / (19*3*3)) 

# Distribution of number of home goals for each team
HG <- test_data %>% 
  group_by(HomeTeam, FTHG) %>%
  summarize(HG = n()) %>%
  spread(FTHG, HG) 

# Distribution of number of away goals for each team
AG <- test_data %>% 
  group_by(AwayTeam, FTAG) %>%
  summarize(AG = n()) %>% 
  spread(FTAG, AG)


```

**Analysis and Parameter Estimation**
```{r}
# First plot the distributions in home and away goals in the same histogram.
(hist_comparison <- ggplot () +
  geom_bar(aes(test_data$FTHG), col = "red", fill = rgb(1, 0, 0, 0.2)) +
  geom_bar(aes(test_data$FTAG), col = "green", fill = rgb(0, 1, 0, 0.2)))

# Intuitively it makes sense that when playing away matches, on a general level, teams tend to apply a more defensive strategy thus scoring less goals. 

# The fit test of poisson returns a favorable result for both home and away goals
gftest_H <- goodfit(test_data$FTHG, type = "poisson", method = "ML")
summary(gftest_H)

gftest_A <- goodfit(test_data$FTAG, type = "poisson", method = "ML")
summary(gftest_A)

# Plot the theoretical poisson distribution and actual home goals
# (hist_h <- ggplot () +
# geom_bar(aes(rpois(1000, gftest_H$par[[1]])), col="black", fill = "gray") +
# geom_bar(aes(test_data$FTHG), col = "red", fill = rgb(1, 0, 0, 0.2)) +
# scale_x_continuous(limits=c(0, 8)) +
# scale_y_continuous(limits=c(0, 15)))
# 
# # Plot the theoretical poisson distribution and actual away goals
# (hist_a <- ggplot () +
# geom_bar(aes(rpois(1000, gftest_A$par[[1]])), col="black", fill = "gray") +
# geom_bar(aes(test_data$FTAG), col = "green", fill = rgb(0, 1, 0, 0.2)))
# # It could be observed that the actual distributions fit with poisson distribution fairly well.

# But not the overall sample is not representative for every team, as strong teams like Man City will still dominate in away matches and score at least two goals most of the times.
test_data %>% 
  group_by(AwayTeam, FTAG) %>%
  filter(AwayTeam ==  "Man City") %>% 
  summarize(FTAG) %>% 
  ggplot() +
  geom_bar(mapping = aes(FTAG))
# Therefore we decided to estimate the distribution parameter for each team

# Summarize goal number for each team in both Home and away matches
HM <- test_data %>% 
  group_by(HomeTeam) %>% 
  summarize(FTHG) %>% 
  pivot_wider(names_from = HomeTeam, values_from = FTHG)

AM <- test_data %>% 
  group_by(AwayTeam) %>% 
  summarize(FTAG) %>% 
  pivot_wider(names_from = AwayTeam, values_from = FTAG)

# Vectorize the fit test function and return parameters for each team for both home and away matches
gftest_vec <- Vectorize(goodfit, "x")
Home_test <- data.frame(t(gftest_vec(HM, "poisson", "ML")))
Away_test <- data.frame(t(gftest_vec(AM, "poisson", "ML")))
pois_par <- data.frame(cbind(unlist(Home_test$par), unlist(Away_test$par)))
colnames(pois_par) = c("Home", "Away")
  
Names <- test_data %>% 
  arrange(HomeTeam) %>%
  select(Team = HomeTeam) %>% 
  unique()  

pois_par <- cbind(Names, pois_par)
pois_par
```

**Monte Carlo Simulation**
```{r}
# Define a function for single-time simulation, with home/away lambda parameters of each team as inputs
sim1 <- function(lambda_data){
  results_h <- data.frame(matrix(NA, nrow = 20, ncol = 20))
  results_a <- data.frame(matrix(NA, nrow = 20, ncol = 20))
  rownames(results_h) = lambda_data[1:20, 1]
  colnames(results_h) = lambda_data[1:20, 1]
  rownames(results_a) = lambda_data[1:20, 1]
  colnames(results_a) = lambda_data[1:20, 1]
  
  for (i in 1:20){
    for (j in 1:20){
      if (i == j){
        results_h[i, j] = 0
      } else {
        if(round(rpois(1, lambda_data[i, 2])) > round(rpois(1, lambda_data[j, 3]))){results_h[i, j] = 3}
        else if(round(rpois(1, lambda_data[i, 2])) < round(rpois(1, lambda_data[j, 3]))){results_h[i, j] = 0}
        else {results_h[i, j] = 1}
      }
    }
  }

  for (i in 1:20){
    for (j in 1:20){
      if (i == j){
        results_a[i, j] = 0
      }
      else if (results_h[j, i] == 3){
        results_a[i, j] = 0
      }
      else if (results_h[j ,i] == 1){
        results_a[i, j] = 1
      }
      else{
        results_a[i, j] = 3
      }
    }
  }
  
  results_sum_h <- results_h %>% 
    rowSums()
  results_sum_a <- results_a %>% 
    rowSums()
  Table <- data.frame(results_sum_h + results_sum_a)
  colnames(Table) <- "Points"
  return(Table)
}

#single_sim <- sim1(pois_par)

# Define Monte Carlo Simulation function.
MCS <- function(n = 1000, lambda_data){
  set.seed(535)
  mc <- data.frame(matrix(NA, nrow = 20, ncol = n))
  rownames(mc) = lambda_data[1:20, 1]
  for (i in 1:n){
    mc[1:20, i] <- sim1(lambda_data)
  }
  
  results <- mc %>% 
    rowSums() %>% 
    as.data.frame()
  
  colnames(results) <- "Points"
  results_mc <- results %>%  
    mutate(Rank = dense_rank(desc(Points))) %>% 
    mutate(Points = round(Points/n, 0)) %>% 
    arrange(Rank)

  return(results_mc)
}

test <- MCS(100, pois_par)

test

# The output makes a lot more sense now, and I believe the next step for us is to figure out the newly promoted team Brentford (never appeared in last 5 seasons in Premier League) in 2021/22 season. There are 30 teams appeared in the last 5 seasons and covered all remaining 19 teams for this season. 
# Please feel free to make visualizations...
```


## Ordinary Least Squares ############################# 

